# Using R

The first implementation of the [pins](https://pins.rstudio.com/) package was made in R.
In this chapter, I will:

 - create a board on my local filesystem.
 - write pins to the board.
 - include the board as a part of this book's website.
 - read pins from the board.

```{r}
#| message: false
library("pins")
library("here")
library("palmerpenguins")
library("waldo")
library("conflicted")
library("pinsManifest") # https://ijlyttle.github.io/pinsManifest/
```

## Local board

The first step is to create a board:

```{r}
board_here <- board_folder(here("pins"), versioned = FALSE)
```

### Writing pins

The next step is to write a pin. Let's write the penguins data-frame as a JSON pin:

```{r}
#| eval: false
pin_write(
  board_here, 
  x = penguins, 
  name = "penguins-json", 
  type = "json",
  metadata = list(
    authors = c("Allison Horst", "Alison Hill", "Kristen Gorman"),
    license = "CCO",
    url = "https://allisonhorst.github.io/palmerpenguins/"
  ),
  versioned = FALSE
)
```

``` default
Creating new version '20220805T171936Z-fa33e'
Writing to pin 'penguins-json'
```

As you can see, the version number is a combination of the creation time (UTC) and a (shortened) hash of the contents.

I also want to create an [arrow](https://arrow.apache.org/docs/r/) version of the pin. 

The `pin_write()` function offers `type = "arrow"`, which uses `arrow::write_feather()`.
However, the default behavior is to use compression; pins does not offer (so far as I know) a way to supply the `compression` argument to `arrow::write_feather()`. 
This presents a problem for me because the arrow implementation for JavaScript does not support compression.

It should not surprise you that pins offers an escape hatch, I can wrap `pins_upload()` in a function:

```{r}
pin_write_arrow_uncompressed <- function(board, x, name = NULL, ...) {
  
  tempfile <- withr::local_tempfile()
  
  arrow::write_feather(x, tempfile, compression = "uncompressed")
  
  result <- pins::pin_upload(
    board,
    paths = tempfile,
    name = name,
    ...
  )
  
  message(glue::glue("Writing to pin '{name}'"))
  
  invisible(result)
}
```

```{r}
#| eval: false
pin_write_arrow_uncompressed(
  board_here, 
  x = penguins, 
  name = "penguins-arrow", 
  metadata = list(
    authors = c("Allison Horst", "Alison Hill", "Kristen Gorman"),
    license = "CCO",
    url = "https://allisonhorst.github.io/palmerpenguins/"
  )
)
```

``` default
Creating new version '20220805T175034Z-ef034'
Writing to pin 'penguins-arrow'
```

### Reading pins 

```{r}
penguins_json <- pin_read(board_here, name = "penguins-json")
compare(penguins, penguins_json)
```

We see some differences between the original ("old") version and "new" version of penguins:

 - new version does not have the "tibble" classes.
 - new version does not know that some of the colunms are factors.
 
These are not huge differences; in fact, the JSON format has no way of encoding that something is a factor.

Let's look at the arrow version.
Because we used a file format (using `pin_upload()`), we need also to write a handler for `pin_download()`:

```{r}
pin_read_arrow_uncompressed <- function(board, name, ...) {
  
  tempfile <- pins::pin_download(board, name, ...)
  
  arrow::read_feather(tempfile)
}
```

```{r}
penguins_arrow <- pin_read_arrow_uncompressed(board_here, "penguins-arrow")
compare(penguins, penguins_arrow)
```

The fact that there are no differences is one of the many cool things about arrow.

### Deploying pins

To make it easier to deploy a board on GitHub (or any other web server), I am using the [pinsManifest](https://ijlyttle.github.io/pinsManifest/) package to create a manifest of pins.
This will make it easier to create a `board_url()` to read pins.

```{r}
#| eval: false
pin_write_manifest(board_here)
```

## Remote board

With this board now availble using GitHub Pages, we can use `board_url()`, which I think is a great way to share data publicly, i.e. without requiring authentication.

```{r}
#| eval: false
board_remote <- 
  board_url_manifest("https://ijlyttle.github.io/pins-three-ways/pins/")
```

### Reading pins

```{r}
#| eval: false
penguins_json_remote <- pin_read(board_remote, name = "penguins-json")
compare(penguins_json, penguins_json_remote)
```
